---
title: "R Notebook"
output: html_notebook
---

### The data science work flow

```{r setup, include = FALSE, warning=FALSE, message=FALSE}

library(tidyverse)
library(tidyquant)
library(timetk)
library(tibbletime)
library(scales)
library(highcharter)
library(broom)
library(PerformanceAnalytics)
knitr::opts_chunk$set(message = FALSE, warning = FALSE, comment = NA)
```


### Introducing R and RStudio

    + Statistical programming language -> by data scientists, for data scientists
    + Base R + 17,000 packages
    + RStudio
    + Shiny
  

### Packages for today


```{r, eval = FALSE}
library(tidyverse)
library(tidyquant)
library(timetk)
library(tibbletime)
library(scales)
library(highcharter)
library(broom)
library(PerformanceAnalytics)
```

More packages for finance here:
https://cran.r-project.org/web/views/Finance.html

### Today's project


    + Buy when the SP500 50-day SMA is above the 200-day SMA (a 'golden cross')
    
    + Sell when the 50-day SMA moves below the 200-day SMA (a 'death cross'). 

    + Code up that strategy

    + Visualize its results and descriptive statistics

    + Compare it to buy-and-hold
    
    + Add secondary logic, visualize that strategy (time permitting)

    + Conclude by building a Shiny dashboard for further exploration
    
    + Data science work flow


### Import data 

We will be working with SP500 and treasury bills data so that when we exit the SP500 we can invest in treasuries and get some return.

We can use the `tidyquant` package and it's `tq_get()` function to grab the data from Yahoo! Finance. 

```{r}

symbols <- c("XLB", "SPY")


prices <- 
  tq_get(symbols, 
         get = "stock.prices",
         from = "1999-01-01") %>% 
  group_by(symbol)

prices  %>% 
  slice(1:3)

prices %>% 
  skimr::skim()
```


Start with a line chart.

We will use `highcharter` to create a quick interactive chart.

```{r}
library(highcharter)
prices %>%
  hchart(., 
         hcaes(x = date, y = adjusted, group = symbol),
         type = "line") %>% 
  hc_title(text = "
           Explore prices")
```


```{r}
returns <- 
prices %>% 
  select(symbol, date, adjusted) %>% 
  mutate(returns = log(adjusted) - log(lag(adjusted)))

```

```{r}
returns %>% 
  tq_performance(Ra = returns,
                 performance_fun = table.Stats) %>% 
  
  # format for easy viewing
  t() %>% 
  knitr::kable()
```


```{r}
returns %>% 
  hchart(., hcaes(x = date, y = returns, group = symbol),
               type = "scatter") %>% 
  hc_tooltip(pointFormat = '{point.x: %Y-%m-%d} <br>
                            {point.y:.4f}%')
```



```{r}
library(readr)
hpi <- 
  read_csv("fmhpi_master_file (2).csv", 
    col_types = cols(GEO_Code = col_skip(), 
        Index_NSA = col_skip())) %>% 
  unite("date", c("Year", "Month"), sep = "-") %>% 
  mutate(date = ymd(parse_date_time(date, "%Y%m")))


hpi %>%
  select(GEO_Type) %>% 
  distinct()
```

```{r}
hpi_for_join <- 
  hpi %>% 
  filter(GEO_Type == "US" & date >= "1999-01-01") %>% 
  rename(housing_index = Index_SA) %>% 
  dplyr::select(-GEO_Name)
```

Looks good, except this is monthly data, and our price data is daily. 

```{r}
xlb_hpi_joined <- 
  returns %>% 
  full_join(hpi_for_join, by = "date")
```

```{r}
xlb_hpi_joined %>% 
  hchart(., hcaes(x = date, y = adjusted, group = symbol),
         type = "line") %>%
  hc_add_series(xlb_hpi_joined, hcaes(x = date, y = housing_index), type = "line")

xlb_hpi_joined %>%
  ggplot(aes(x = date, y = adjusted, color = symbol)) +
  geom_line() +
  geom_line(aes(y = housing_index, color = GEO_Type))
```

```{r}
prices %>% 
  select(date, symbol, adjusted) %>%
  as_tbl_time(index = date) %>% 
  as_period("monthly", side = "start") %>%
  mutate(date = floor_date(date, "month")) %>% 
  full_join(hpi_for_join) %>% 
  mutate(hpi_change = log(housing_index) - log(lag(housing_index))) %>% 
  ggplot(aes(x = date, y = adjusted, color = symbol)) +
  geom_line() +
  geom_line(aes(y = housing_index, color = GEO_Type)) +
    facet_wrap(~symbol)
```

```{r}
xlb_hpi_joined_monthly <-
prices %>% 
  select(date, symbol, adjusted) %>%
  as_tbl_time(index = date) %>% 
  as_period("monthly", side = "start") %>%
  mutate(date = floor_date(date, "month")) %>% 
  full_join(hpi_for_join) %>% 
  mutate(hpi_change = log(housing_index) - log(lag(housing_index)),
         asset_returns = log(adjusted) - log(lag(adjusted))) %>% 
  na.omit()
  
xlb_hpi_joined_monthly %>% 
  summarise(mean_house_change = mean(hpi_change))
```

### Add some trading logic

If the previous two HPI monthly changes were below mean, exit the XLB position.

```{r}
xlb_hpi_joined_monthly %>%
  filter(symbol == "XLB") %>% 
  mutate(signal = if_else(
                            lag(hpi_change, 1) < mean(hpi_change) &
                            lag(hpi_change, 2) < mean(hpi_change) &
                            lag(hpi_change, 3) < mean(hpi_change),
                            0, 1),
         trend_returns = if_else(lag(signal) == 1, 
                                 (signal * asset_returns), 0),
         buy_hold_returns = (asset_returns)) %>% 
  select(date, symbol, buy_hold_returns, trend_returns) %>%
  na.omit() %>% 
  mutate(
         trend_growth = accumulate(1 + trend_returns, `*`),
         buy_hold_growth = accumulate(1 + buy_hold_returns, `*`))
  
```

```{r}
xlb_hpi_joined_monthly %>%
  filter(symbol == "SPY") %>% 
  mutate(signal = if_else(
                             lag(hpi_change, 1) < 0, #mean(hpi_change), #&
                            # lag(hpi_change, 2) < mean(hpi_change) &
                            # lag(hpi_change, 3) < mean(hpi_change) &
                            # lag(hpi_change, 4) < mean(hpi_change),
                            0, 1),
         trend_returns = if_else(lag(signal) == 1, 
                                 (signal * asset_returns), 0),
         buy_hold_returns = (asset_returns)) %>% 
  select(date, symbol, buy_hold_returns, trend_returns) %>%
  na.omit() %>% 
  mutate(
         trend_growth = accumulate(1 + trend_returns, `*`),
         buy_hold_growth = accumulate(1 + buy_hold_returns, `*`)) %>%
  select(date, symbol, trend_growth, buy_hold_growth) %>%
  gather(strategy, growth, -date, -symbol) %>% 
  #filter(symbol == "XLB") %>% 
  hchart(., hcaes(x = date, y = growth, group = strategy), type = "line") %>% 
  hc_tooltip(pointFormat = "{point.strategy}: ${point.growth: .2f}")
  
```

